\section{Attention Residuals: A Unified View of Time and Depth}
\label{sec:attnres}

The limitations discussed above are reminiscent of similar bottlenecks in sequence modeling, suggesting that we seek similar solutions for the depth dimension.

\paragraph{The Duality of Time and Depth.}
Like RNNs over time, residual connections compress all prior information into a single state $\bm{h}_l$ over depth.
For sequence modeling, the Transformer improved upon RNNs by replacing recurrence with attention~\citep{bahdanau2016attention,vaswani-2017-attention}, allowing each position to selectively access all previous positions with data-dependent weights.
We propose the same methodology for depth:
\begin{equation}
    \bm{h}_{l} = \brickred{\alpha_{0 \to l}} \cdot \bm{h}_1 + \sum_{i=1}^{l-1} \brickred{\alpha_{i \to l}} \cdot f_i(\bm{h}_{i})
\end{equation}
where $\brickred{\alpha_{i \to l}}$ are layer-specific attention weights satisfying $\sum_{i=0}^{l-1} \brickred{\alpha_{i \to l}} = 1$.
Unlike sequence length (which can reach millions of tokens), network depth is typically modest ($L < 1000$), making $O(L^2)$ attention over depth computationally feasible.
We call this approach \emph{Attention Residuals}, abbreviated as \emph{AttnRes}.

\subsection{Full Attention Residuals}
The attention weights can be written as $\brickred{\alpha_{i \to l}} = \phi(\bm{q}_{l},\, \bm{k}_{i})$ for a kernel function $\phi\colon \mathbb{R}^d \times \mathbb{R}^d \to \mathbb{R}_{\geq 0}$, where $\bm{q}_l$ and $\bm{k}_i$ are query and key vectors~\cite{katharopoulos-2020-transformers, zhong2025understanding}. Different choices of $\phi$ recover different residual variants (\S\ref{sec:structured-depth}); we adopt $\phi(\bm{q}, \bm{k}) = \exp\left(\bm{q}^\top\operatorname{RMSNorm}(\bm{k})\right)$~\cite{zhang2019root} with normalization, yielding $\operatorname{softmax}$ attention over depth:
\begin{equation}
    \label{eq:attn-weights}
    \brickred{\alpha_{i \to l}} = \frac{\phi\left(\bm{q}_{l}, \bm{k}_{i}\right)}{\sum_{j=0}^{l-1} \phi\left(\bm{q}_{l}, \bm{k}_{j}\right)}
\end{equation}

For each layer $l$, we define:
\begin{equation}
    \label{eq:qkvattn}
    \bm{q}_{l} = \bm{w}_{l}, \quad
    \quad \bm{k}_{i} = \bm{v}_{i} = \begin{cases} \bm{h}_1 & i = 0 \\ f_i(\bm{h}_{i}) & 1 \leq i \leq l-1 \end{cases}
\end{equation}
where the query $\bm{q}_l=\bm{w}_{l}$ is a layer-specific learnable vector in $\mathbb{R}^{d}$. The $\operatorname{RMSNorm}$ inside $\phi$ prevents layers with large-magnitude outputs from dominating the attention weights. The input to layer $l$ is then:
\begin{equation}
    \label{eq:full-attn}
    \bm{h}_{l} = \sum_{i=0}^{l-1} \brickred{\alpha_{i \to l}} \cdot \bm{v}_{i}
\end{equation}





We call this form \emph{full attention residuals}. For each token, Full AttnRes requires $O(L^2 d)$ arithmetic and $O(Ld)$ memory to store layer outputs. Since depth is far smaller than sequence length, the arithmetic cost is modest.

\paragraph{Overhead.}
The $O(Ld)$ memory overlaps entirely with the activations already retained for backpropagation, so Full AttnRes introduces no additional memory overhead in vanilla training. At scale, however, activation recomputation and pipeline parallelism are widely adopted: layer outputs that would otherwise be freed and recomputed must now be kept alive for all subsequent layers, and under pipeline parallelism each must further be transmitted across stage boundaries. Both the memory and communication overhead then grow as $O(Ld)$.

\paragraph{Blockwise optimization.}
A deliberate design choice in Full AttnRes is that the \emph{pseudo}-query $\bm{w}_l$ is a learned parameter decoupled from the layer's forward computation. This independence means that attention weights for any group of layers can be computed in parallel without waiting for their sequential outputs, and in particular permits grouping the $L$ layers into $N$ blocks of $S$ layers each and batching the attention computation within each block, reducing per-layer memory I/O from $O(Ld)$ to $O((S{+}N)d)$ (we defer the detailed two-phase strategy to \S\ref{sec:infra}). Under current distributed training regimes, however, the dominant cost is not local memory bandwidth but cross-stage communication under pipeline parallelism: every layer output must still be transmitted between stages, and this $O(Ld)$ communication overhead cannot be alleviated by local batching. This motivates the Block AttnRes variant introduced below, which reduces the number of cross-stage representations from $L$ to $N$. We anticipate that future interconnect improvements will make the full $O(Ld)$ communication practical, fully realizing the potential of Full AttnRes.

\subsection{Block Attention Residuals}

We propose \emph{Block Attention Residuals}, which partitions the $L$ layers into $N$ blocks: within each block, the layer outputs are reduced to a single representation via summation, and across blocks, we apply full attention over only $N$ block-level representations and the token embedding. This reduces both memory and communication overhead from $O(Ld)$ to $O(Nd)$.

\paragraph{Intra-Block Accumulation.}
Specifically, we divide the $L$ layers into $N$ blocks of $S = L/N$ layers each, assuming $L$ is divisible by $N$; otherwise, the last block contains the remaining $L \bmod N$ layers. Let $\mathcal{B}_n$ denote the set of layer indices in block $n$ ($n = 1, \ldots, N$).
To form a block, we sum all of its layer outputs:
\begin{equation}
    \bm{b}_n = \sum_{j \in \mathcal{B}_n} f_j(\bm{h}_j)
\end{equation}
We further denote $\bm{b}_n^i$ as the partial sum over the first $i$ layers in $\mathcal{B}_n$, so that $\bm{b}_n = \bm{b}_n^{S}$.
When $L$ is not divisible by $N$, the final partial sum is taken as the last block's representation.
As in Full AttnRes, the $\operatorname{RMSNorm}$ inside $\phi$  prevents magnitude differences between complete blocks and partial sums from biasing the attention weights.

\paragraph{Inter-Block Attention.}
In Full AttnRes, the input to layer $l$ is computed by attending over all outputs up to $f_{l-1}(\bm{h}_{l-1})$.
The block-wise variant replaces these individual outputs with block representations, defining $\bm{b}_0 = \bm{h}_1$ so that the token embedding is always included as a source.
For the $i$-th layer in block $n$, the value matrix is:
\begin{equation}
    \mathbf{V} = \begin{cases}
        [\bm{b}_0, \bm{b}_1, \ldots, \bm{b}_{n-1}]^\top                 & \text{if } i = 1 \text{ (first layer of block } n\text{)} \\
        [\bm{b}_0, \bm{b}_1, \ldots, \bm{b}_{n-1}, \bm{b}_n^{i-1}]^\top & \text{if } i \geq 2 \text{ (subsequent layers)}           \\
    \end{cases}
\end{equation}
Keys and attention weights follow Eq.~\ref{eq:qkvattn} and Eq.~\ref{eq:attn-weights}.
The input of the very first layer of the network is the token embeddings, i.e. $\bm{b}_0 = \bm{h}_1$. In each block, the first layer receives the previous block representations and the token embeddings, and the subsequent layers additionally attend to the partial sum $\bm{b}_n^{i-1}$. The final output layer aggregates all $N$ block representations.
Fig.~\ref{listing:attnres} provides PyTorch-style pseudocode for Block AttnRes.
\input{figures/code}

\paragraph{Efficiency.}
Since each layer now attends over $N$ block representations rather than $L$ individual outputs, memory reduces from $O(L)$ to $O(N)$ and computation from $O(L^2)$ to $O(N^2)$.
The block count $N$ interpolates between two extremes: $N = L$ recovers Full AttnRes, while $N = 1$ reduces to standard residual connections with the embedding isolated as $\bm{b}_0$.
Empirically, we find that $N \approx 8$ recovers most of the benefit across model scales, requiring only eight stored hidden states per token (see \S~\ref{sec:exp}).

Beyond memory and computation, the block structure also benefits inference latency: block boundaries define the dispatch granularity for the blockwise optimization described in \S\ref{sec:attnres}, and the fixed block count $N$ bounds the KV cache size. The parallel inter-block results are merged with the sequential intra-block partial sums via online $\operatorname{softmax}$~\citep{milakov2018online}, preserving exact equivalence (\S\ref{sec:infra}).